---
title: "datafile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{datafile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}

install.packages("remotes", repos='https://mirror.lyrahosting.com/CRAN/')
remotes::install_github("jspijker/datafile")
```

# Introduction

The datafile package is build around the philosophy that data should
be separated from code. This means that data is always written to a
different location (directory) than your current working directory,
which contains the code.

To save you from the hassle of creating a data directory or to type
the data directory filepath over and over again, the datafile package
contains some convienient functions. 

Instead of typing filepaths, you indicate where data is stored and
then use the `datafile()` function to expand your filename into a full
file path. So the result of datafile("somedata.rds") can be
'/home/user/work/project/data/somedata.rds'.

The location of the data directory depends on how you initialize
datafile. It is aimed to be flexible so it can suit many different
workflows of scenario's.

To illustrate the datafile package we give a few examples and usage
scenarios.

# Examples

## Example 1, basic usage

Let's start with a basic example:

```{r,eval=FALSE}
library(datafile)
datafileInit() 
x<-rnorm(10)
saveRDS(x,datafile("x.rds"))

```

In this example we create a data directory using `datafileInit()` under
the current root directory of your project. Datafileinit uses the
here() function to obtain this root directory. Under this root
directory, a directory named 'data' is created.

Then we created some variable `x`. And
using saveRDS and datafile we save it to disk. The final RDS file is
now saved to `./data/x.rds`, relative to the project root directory.

It is good practice to create data directories as relative paths to you current
working directory or project root, e.g. ../../data or ./data. If you use absolute
paths, e.g. /home/johndoe/project/data, your code won't be reproducible
by other users. 

To retrieve the rds file, we use:

```{r,eval=FALSE}
x<-readRDS(datafile("x.rds"))
```

## Example 2, different base directory

Now let's assume you want to store your data in a temporary directory,
because you are working using containers and only use intermediate
data. In that case you can provide datafileInit with a specific base
directory, absolute or relative. The above example then looks like this:


```{r}
datafileInit(basedir=tempdir()) 
x<-rnorm(10)
saveRDS(x,datafile("x.rds"))

# read it back:
x<-readRDS(datafile("x.rds"))
```

The only things that is changed, is the location of the base
directory. Datafile expands your filenames to the right directory.

## Example 3, using datadir argument

One other possibility to use datafile is for example when you want to
store different itterations of your data. By using the datadir argument
you can change the name of the data directory. Let's take the same
example as above and use the datadir argument.

```{r}
datafileInit(basedir=tempdir(),datadir='data01') 
x<-rnorm(10)
saveRDS(x,datafile("x.rds"))

# read it back:
x<-readRDS(datafile("x.rds"))
```

Now if you want to create a new itteration of your model and want to
store the new data somewhere else, just change the datadir directory.
Please not that all other code isn't changed.

```{r}
datafileInit(basedir=tempdir(),datadir='data02') 
x<-rnorm(10)
saveRDS(x,datafile("x.rds"))
```

## Example 4, using the subdir argument

If you use git for your data science projects, the default basedir for
datafile is set at the root of the git repository. Hence, the default data
directory will be at the root of your git repository. 

One way of managing your project is to use diferent subdirectories for
the different steps in your project, for example data import, analyses
and reporting. With datafile you can create separate data directories
in each subdirectory using the subdir argument. 

```{r eval=FALSE}
datafileInit(subdir='import') 
x<-rnorm(10)
saveRDS(x,datafile("x.rds"))

# read it back:
x<-readRDS(datafile("x.rds"))
```

In this case the data is stored at './import/data/x.rds'


## Example 5, short notation

There is one more thing! If `datafile` is to much typing for you, you can use the alias
`dfn()`:

```{r,eval=FALSE}
x<-readRDS(dfn("x.rds"))

```


```{r}
#unlink("./data/",recursive=TRUE)
```

